# Predict what makes a developer more likely to work remotely

Stack Overflow is the world's largest, most trusted online community for developers (I bet you have used it!) and every year there is a large survey of developers, to learn about developers' opinions on different technologies, work habits, and so forth.
You can find the link to the surveys here: 

:::{.callout-note}
## Learning objective:
- You are going to use data from the annual Developer Survey to build predictive models;
- You'll do exploratory data analysis to understand what's in the dataset and how some of the quantities in the survey are distributed;
-You'll practice your machine learning skills by training classification/regression models.
:::

:::{.callout-tip}
## Exercise:

In this case study, you will predict whether a developer works remotely or not (i.e., in their company offices) from characteristics of these developers, like experience and size of the company. In this analysis, we will assume that a software developer can either work remotely, or not. What kind of model will you build?
:::

Load the libraries:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```

Load the data:


```{r}
library(readr)
StackSurvey <- read_csv("../datasets/stack_overflow.csv", show_col_types = FALSE)
```

Start off this modeling analysis by checking out how many remote and non-remote developers you have to work with, where they live, and how much experience they have:

```{r}
# Take a look at StackSurvey
glimpse(StackSurvey)

# First count for `remote`
StackSurvey %>% 
    count(remote, sort = TRUE)

# then count for `country`
StackSurvey %>% 
    count(country, sort = TRUE)
```

Use the appropriate column from the data set so you can plot a boxplot with remote status on the x-axis and professional experience on the y-axis:

``` {r}
ggplot(StackSurvey, aes(remote, years_coded_job)) +
    geom_boxplot() +
    labs(x = NULL,
         y = "Years of professional coding experience")
```
